---
title: Ribs (Bone)
---

# Overview

There are 12 pairs of rib bones which correspond to the 12 thoracic vertebrae (T1-T12).
These ribs structurally function to protect the vital organs but also serve as functional levers for respiration.


# Articulations

Most of the ribs have a *head*, *neck*, and *articular tubercle* posteriorly which articulate with the [costovertebral joints]({{< var ref-costovertebral-joint.path >}}) which consist of a [costocorporeal joint]({{< var ref-costocorporeal-joint.path >}}) and [costotransverse joint]({{< var ref-costotransverse-joint.path >}}) @neumannKinesiologyMusculoskeletalSystem2017.

The [costocorporeal joint]({{< var ref-costocorporeal-joint.path >}}) connects the [*head*](#head) of the rib to the *costal demifacets* of the two corresponding vertebrae and the [intervertebral disc]({{< var ref-intervertebral-disc.path >}}) @neumannKinesiologyMusculoskeletalSystem2017.

The [costotransverse joint]({{< var ref-costotransverse-joint.path >}}) refers to the articulation between the *articular tubercle* of the rib with the *costal facet* of the *transverse process* of the corresponding vertebrae @neumannKinesiologyMusculoskeletalSystem2017

## Sternocostal joints



# Function



# Anatomy

The anterior aspect of the rib is composed of flattened *hyaline cartilage* @neumannKinesiologyMusculoskeletalSystem2017.
Ribs 1-10 span from the spine posteriorly and span anteriorly to directly or indirectly connect to the [sternum]({{< var ref-sternum-bone.path >}}) to complete the rib cage @neumannKinesiologyMusculoskeletalSystem2017.
Ribs 1-7 articulate *directly* with the sternum via [sternocostal joints]({{< var ref-sternocostal-joint.path >}}) @neumannKinesiologyMusculoskeletalSystem2017.

## Head {#head}

The head of the rib makes up the distal aspect of the [costocorporeal joint]({{< var ref-costocorporeal-joint.path >}})

## Neck {#neck}

## Articular tubercle {#articular-tubercle}


# Muscular attachments

```{r tbl-rib-origin-oian}
#| echo: false
#| warning: false
#| message: false

# Load the here package
library(here)

# Source the functions script
source(here("scripts/gt_oian.R"))

# Example usage
identifier_col <- "origin_identifier"  # or "group_identifier"
filter_patterns <- "1st_rib_bone | 2nd_rib_bone | 3rd_rib_bone | 4th_rib_bone | 5th_rib_bone | 6th_rib_bone | 7th_rib_bone | 8th_rib_bone | 9th_rib_bone | 10th_rib_bone | 11th_rib_bone | 12_rib_bone"
display_columns <- c("name_final", "origin_final", "insertion_final", "innervation_final", "action_final")
table_title <- "Muscles originating from the ribs"

# Call the function
gt_table <- gt_oian(identifier_col, filter_patterns, display_columns, table_title)

gt_table
```


```{r tbl-rib-insertion-oian}
#| echo: false
#| warning: false
#| message: false

# Load the here package
library(here)

# Source the functions script
source(here("scripts/gt_oian.R"))

# Example usage
identifier_col <- "insertion_identifier"  # or "group_identifier"
filter_patterns <- "1st_rib_bone | 2nd_rib_bone | 3rd_rib_bone | 4th_rib_bone | 5th_rib_bone | 6th_rib_bone | 7th_rib_bone | 8th_rib_bone | 9th_rib_bone | 10th_rib_bone | 11th_rib_bone | 12_rib_bone"
display_columns <- c("name_final", "origin_final", "insertion_final", "innervation_final", "action_final")
table_title <- "Muscles inserting on the ribs"

# Call the function
gt_table <- gt_oian(identifier_col, filter_patterns, display_columns, table_title)

gt_table
```


# Ribs

## 1st

[First rib]({{< var ref-1st-rib.path >}})

## 2nd

## 3rd

## 4th

## 5th

## 6th

## 7th

## 8th

## 9th

## 10th

## 11th 

"Free floating rib"

## 12th

The 12th rib is considered a "Free floating rib" since it lacks attachment to the sternum