---
title: Sacrotuberous ligament
---




Glute max partly originates from this ligament


# Myofascial unit

The *sacrotuberous lig.* plays a dual role.
The deeper fibers attach bone-to-bone and stabilize the SIJ.
The superficial fibers are important in creating myofascial lines @myersAnatomyTrains2021.

If there is too much movement in the *deeper fibers* this would be considered a lax ligament and result in SIJ instability @myersAnatomyTrains2021.
Too *little* movement in the superficial region will result in rigidity and immobility in the myofascial lines, reducing the body's resilience in movement @myersAnatomyTrains2021.
# Muscular Attachment

```{r}
#| echo: false
#| warning: false
#| message: false

# Load the here package
library(here)

# Source the functions script
source(here("scripts/gt_oian.R"))

# Example usage
identifier_col <- "origin_identifier"  # or "group_identifier"
filter_patterns <- "sacrotuberous_ligament"
display_columns <- c("name_final", "origin_final", "insertion_final", "innervation_final", "action_final")
table_title <- "Muscles originating from the Sacrotuberous ligament"

# Call the function
gt_table <- gt_oian(identifier_col, filter_patterns, display_columns, table_title)

gt_table
```

```{r}
#| echo: false
#| warning: false
#| message: false

# Load the here package
library(here)

# Source the functions script
source(here("scripts/gt_oian.R"))

# Example usage
identifier_col <- "insertion_identifier"  # or "group_identifier"
filter_patterns <- "sacrotuberous_ligament"
display_columns <- c("name_final", "origin_final", "insertion_final", "innervation_final", "action_final")
table_title <- "Muscles Inserting onto the Sacrotuberous ligament"

# Call the function
gt_table <- gt_oian(identifier_col, filter_patterns, display_columns, table_title)

gt_table
```